\documentclass[12pt,letterpaper]{article}
\usepackage[latin1]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{amsthm}

\newtheorem*{eg}{Example}

\author{Linyun Fu}
\title{CSCI 4020 Computer Algorithms Spring 2011\\
Problem Set 8}
\begin{document}
\maketitle
\section*{Chapter 8, Problem 3}
Suppose you're helping to organize a summer sports camp, and the
following problem comes up. The camp is supposed to have at least
one counselor who's skilled at each of the n sports covered by the camp
(baseball, volleyball, and so on). They have received job applications from
$m$ potential counselors. For each of the $n$ sports, there is some subset
of the $m$ applicants qualified in that sport. The question is: For a given
number $k < m$, is it possible to hire at most $k$ of the counselors and have
at least one counselor qualified in each of the $n$ sports? We'll call this the
{\em Efficient Recruiting Problem}.

Show that Efficient Recruiting is NP-complete.

\section*{Answer}

\section*{Chapter 8, Problem 9}
Consider the following problem. You are managing a communication
network, modeled by a directed graph $G = (V, E)$. There are $c$ {\em users} who
are interested in malting use of this network. User $i$ (for each $i = 1, 2, \dots, c$)
issues a {\em request} to reserve a specific path $P_i$ in $G$ on which to transmit
data.

You are interested in accepting as many of these path requests as
possible, subject to the following restriction: if you accept both $P_i$ and $P_j$,
then $P_i$ and $P_j$ cannot share any nodes.

Thus, the {\em Path Selection Problem} asks: Given a directed graph $G =
(V, E)$, a set of requests $P_1, P_2, \dots, P_c$ -- each of which must be a path in
$G$ -- and a number $k$, is it possible to select at least $k$ of the paths so that
no two of the selected paths share any nodes?

Prove that Path Selection is NP-complete.

\section*{Answer}

\end{document}

