\documentclass[12pt,letterpaper]{article}
\usepackage[latin1]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{amsthm}

\newtheorem*{eg}{Example}

\author{Linyun Fu}
\title{CSCI 4020 Computer Algorithms Spring 2011\\
Problem Set 7}
\begin{document}
\maketitle
\section*{Chapter 7, Problem 29}
Some of your friends have recently graduated and started a small company,
which they are currently running out of their parents' garages in
Santa Clara. They're in the process of porting all their software from an
old system to a new, revved-up system; and they're facing the following
problem.

They have a collection of $n$ software applications, $\{1, 2, \dots, n\}$, running
on their old system; and they'd like to port some of these to the new
system. If they move application $i$ to the new system, they expect a net
(monetary) benefit of $b_i \ge 0$. The different software applications interact
with one another; if applications $i$ and $j$ have extensive interaction, then
the company, will incur an expense if they move one of $i$ or $j$ to the new
system but not both; let's denote this expense by $x_{ij} \ge 0$.

So, if the situation were really this simple, your friends would just
port all $n$ applications, achieving a total benefit of $\sum_i b_i$. Unfortunately,
there's a problem, ...

Due to small but fundamental incompatibilities between the two
systems, there's no way to port application 1 to the new system; it will
have to remain on the old system. Nevertheless, it might still pay off to
port some of the other applications, accruing the associated benefit and
incurring the expense of the interaction between applications on different
systems.

So this is the question they pose to you: Which of the remaining
applications, if any, should be moved? Give a polynomial-time algorithm
to find a set $S \subseteq \{2, 3, \dots, n\}$ for which the sum of the benefits minus the
expenses of moving the applications in $S$ to the new system is maximized.

\section*{Answer}

\section*{Chapter 7, Problem 30}
Consider a variation on the previous problem. In the new scenario, any
application can potentially be moved, but now some of the benefits $b_i$ for
moving to the new system are in fact negative: If $b_i < 0$, then it is preferable
(by an amount quantified in $b_i$) to keep $i$ on the old system. Again, give
a polynomial-time algorithm to find a set $S \subseteq \{1, 2, \dots, n\}$ for which the
sum of the benefits minus the expenses of moving the applications in $S$
to the new system is maximized.

\section*{Answer}

\end{document}
